Data cleaning and processing script

### 2023-02-23 code/processing_code/data_processing_LML.qmd
exploratory analysis

@@

#Load Catchment-level COVID-19 Case Data
#covid_case_wrf = read_csv(here("data/raw_data/cases_wrf.csv"))
```

@@

1.  CFX N1 results
2.  CFX N2 results
3.  Plant flow and total suspended solids
4.  County-level COVID symptoms
5.  County-level COVID confirmed cases
6.  County-level COVID tests administered
7.  Catchment-level COVID confirmed cases

@@

wbe_year2 <- bind_rows(n1, n2) %>% #combine dataframes with same number of rows
  mutate(
    date=as.Date(sample_date, format = "%d-%b-%y"), #change to date format
    facility=substr(sample_id, 1,2), #first two letters in sample_id is facility ID
    biological_replicate=substr(sample_id, nchar(sample_id), nchar(sample_id)), #last number in sample_id is the replicate number
    ct=as.numeric(ifelse(ct=="Undetermined", NA, ct)) #set NAs
  ) %>% 
  arrange(date, facility, target, biological_replicate) %>% 
  select(date, facility, target, biological_replicate, collection_num, ct, copy_num_uL_rxn) %>%
  mutate(facility = as.factor(facility), facility = recode(facility, NO = "A", MI = "B", CC = "C"), facility = ordered(facility, levels = c("A", "B", "C"))) %>% #set facility to factor and then set factor levels
  filter(between(date, as.Date("2021-07-06") , as.Date("2022-12-21"))) 
```

@@

@@

```{r}
n1_n2_plant <- left_join(wbe_year2, plant_clean, by = c("date","facility"))
```

```{r}
covid <- full_join(
  covid%>%
    select(cases.symptom.onset=cases, date=symptom.date), 
  covid.report%>%
    select(cases.reported=cases, date=report_date), 
  by = "date"
) %>% 
  full_join(
    covid.testing%>%
      rename(date=collection_date), 
    by="date"
  ) %>%
  select(date, cases.symptom.onset, cases.reported, pcr_tests, pcr_pos, pcr_pos_7dma, pcr_pos_14dma) 
  #full_join(cases.wrf, by = "date")
```

```{r}
all_data <- full_join(n1_n2_plant, covid, by = "date")
```

>>>>

recovery_output \<- read_csv(here("data/raw_data/recovery_data.csv"))

recovery_input \<- read_csv(here("data/raw_data/calfguard.csv"))

@@

#Load Catchment-level COVID-19 Case Data
#covid_case_wrf = read_csv(here("data/raw_data/cases_wrf.csv"))

#Load CFX recovery data
recovery_output <- read_csv(here("data/raw_data/recovery_data.csv"))
recovery_input <- read_csv(here("data/raw_data/calfguard.csv"))
```

@@

1.  CFX N1 results
2.  CFX N2 results
3.  Plant flow and total suspended solids
4.  County-level COVID symptoms
5.  County-level COVID confirmed cases
6.  County-level COVID tests administered
7.  Catchment-level COVID confirmed cases
8.  CFX recovery input
9.  CFX recovery output

@@

wbe_year2 <- bind_rows(n1, n2) %>% #combine dataframes with same number of rows
  mutate(
    date=as.Date(sample_date, format = "%d-%b-%y"), #change to date format
    facility=substr(sample_id, 1,2), #first two letters in sample_id is facility ID
    biological_replicate=substr(sample_id, nchar(sample_id), nchar(sample_id)), #last number in sample_id is the replicate number
    ct=as.numeric(ifelse(ct=="Undetermined", NA, ct)) #set NAs
  ) %>% 
  arrange(date, facility, target, biological_replicate) %>% 
  select(date, facility, target, biological_replicate, collection_num, ct, copy_num_uL_rxn) %>%
  mutate(facility = as.factor(facility), 
         facility = recode(facility, NO = "A", MI = "B", CC = "C"), 
         facility = ordered(facility, levels = c("A", "B", "C")), #set facility to factor and then set factor levels
         copy_num_L = copy_num_uL_rxn*20/5*60/280*1000*1000) %>% 
  filter(between(date, as.Date("2021-07-06") , as.Date("2022-12-21")))

saveRDS(wbe_year2, here("data/processed_data/year2_cfx_clean.rds"))
```

@@

saveRDS(plant_clean, here("data/processed_data/year2_plant_clean.rds"))
```

```{r}
n1_n2_plant <- left_join(wbe_year2, plant_clean, by = c("date","facility")) %>% 
  mutate(total_copies = copy_num_L * influent_flow_L)

saveRDS(n1_n2_plant, here("data/processed_data/year2_cfx_plant_combined.rds"))
```

```{r}
covid <- full_join(
  covid_symptom%>%
    select(cases.symptom.onset=cases, date=symptom.date), 
  covid_case%>%
    select(cases.reported=cases, date=report_date), 
  by = "date"
) %>% 
  full_join(
    covid_testing%>%
      rename(date=collection_date), 
    by="date"
  ) %>%
  select(date, cases.symptom.onset, cases.reported, pcr_tests, pcr_pos, pcr_pos_7dma, pcr_pos_14dma) 
  #full_join(cases.wrf, by = "date")

saveRDS(covid, here("data/processed_data/all_covid_combined.rds"))
```

```{r}
all_data <- full_join(n1_n2_plant, covid, by = "date") %>% 
   filter(between(date, as.Date("2021-07-06") , as.Date("2022-12-21")))

saveRDS(all_data, here("data/processed_data/year2_all_data_combined.rds"))
```

### Cleaning recovery data sets

```{r}
#Start with recovery_output 
recovery_output_clean <- recovery_output %>% 
  mutate(ct = as.numeric(ct)) %>% 
  group_by(sample_id, cg_num) %>% 
  summarize(avg_ct = mean(ct)) %>% 
  separate(sample_id, into = c("facility", "collection_num", "rep"), sep = "_") %>% 
  mutate(collection_num = as.numeric(collection_num),
         copies_ul_rxn = 10^((avg_ct-30.7)/-3.238),
         copies_ul_sample = copies_ul_rxn *20/2*25/3*60/280) %>% 
  select(facility, collection_num, cg_num, copies_ul_sample, avg_ct) %>% 
  drop_na()

recovery_input_clean <- recovery_input %>% 
  mutate(ct = as.numeric(ct)) %>% 
  group_by(sample_id) %>% 
  summarize(avg_ct = mean(ct)) %>% 
  separate(sample_id, into = c("CG", "cg_num"), sep = "_") %>% 
  mutate(cg_num = as.numeric(cg_num),
         copies_ul_rxn = 10^((avg_ct-30.7)/-3.238),
         copies_ul_input = copies_ul_rxn *20/2*25/3*60/50*40/40000) %>% 
  select(cg_num, copies_ul_input, avg_ct) %>% 
  drop_na()

saveRDS(recovery_output_clean, here("data/processed_data/recovery_output_clean.rds"))
saveRDS(recovery_input_clean, here("data/processed_data/recovery_input_clean.rds"))

@@

### Combining recovery data sets

```{r}
recovery_calc <- left_join(recovery_output_clean, recovery_input_clean, by = "cg_num") %>% 
  mutate(percent_recovery = (copies_ul_sample/copies_ul_input)*100) %>% 
  select(facility, collection_num, percent_recovery)

saveRDS(recovery_calc, here("data/processed_data/recovery_percentages.rds"))
```
